---
title: "Penguin Flipper Length ANOVA"
author: "Priscilla Hare"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Load packages and read in penguins.csv
Packages: tidyverse, car, onewaytets
```{r}
#load packages and read in penguins.csv data

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA...
Assuptions:
- Independent observations
-equal variance (Levene's test)
-normality (histograms and qq)


```{r}

counts <- with(penguins, table(species,sex)) #creates a table of counts of species of each sex 

penguins2 <- penguins %>%
  mutate(
    common_name = case_when(
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo"
    )
  ) %>% 
  select(-species) %>% #removes species column
  filter(sex == "FEMALE")

```

Exploratory graph and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name, scale = "free")

hists

qq <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~common_name, scale = "free")

qq

# Based on qq plots and histograms, observations for each group appear normally distrubuted, also n>30 for all groups so means will be normally distributed by central limit theorem

```

Evaluate variances and perform a Levene's test for equal variances
```{r}
# If the largest sample variances is < 4x greater than the smallest samople variance, then these are close enough to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances #the largest variance (33.1) is definitely not more than 4x greater than smallest variacne (15.2)

#formal hypothesis test for equal variances - Levenes test
# leveneTest() function in the 'car' package
levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)

levene_penguin

#H0: no difference in variacnes
#HA: difference in variances

#This would make us conclude that there is no significant in variances across species

```

One-way ANOVA for penguin flipper lengths ny species (common_name)
```{r}

#H0: means across all groups are equal
#HA: at least 2 means differ

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

#Post-hoc testing by Tukey's HSD
#H0: there is no significant difference in pair wise means
#HA: there is a significant difference in pair wise means
penguin_post <- TukeyHSD(penguin_aov)
penguin_post

```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p <0.001, $\alpha$ = 0.05 with a post-hoc Tukey's HSD, $\alpha$ = 0.05)

#cannot do inline referencing with aov